---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "8/22/2022"
output:
  pdf_document: default
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
print(paste("There are total of ",dim(mtcars)[2] ,"variables and",dim(mtcars)[1],"observations in this data set."))
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)
# Answer:
print("There are 5 discrete variables and 6 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m<-mean(mtcars$mpg)
v<-var(mtcars$mpg)
s<-sd(mtcars$mpg)

print(paste("The average of Mile Per Gallon from this data set is ",m , " with variance ",s, " and standard deviation",s, "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
library(dplyr)
table.mean=mtcars %>% group_by(cyl) %>% summarise(mean = mean(mpg))
table.sd=mtcars%>%group_by(cyl) %>% summarise(standard_deviation=sd(mpg))
table.mean
table.sd

```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
xtabs(~cyl+gear,data=mtcars)
#table(mtcars$cyl,mtcars$gear)
print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
library(ggplot2)
graph<-ggplot(PlantGrowth,aes(x=group,y=weight,color=group)) +geom_boxplot()
graph

```

Result:

=> Group trt2 seems to have the largest weight, while group trt1 seems to have the smallest weight.



b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}

hist(mtcars$mpg,breaks=10,main="Histogram of the Mile Per Gallon for cars",xlab="Mile Per Gallon",ylab="Frequency",col="green")

print("Most of the cars in this data set are in the class of 15 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
plot(USArrests$Murder,USArrests$Assault,main = "Relationship between the murder and the assult from all the US arrest",xlab="Murder",ylab="Assault",col="green",pch=19)

```

Result:

Based from the graph, there is a postitive linear relationship between the Murder and the Assault in the US arrest.


***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
housingData.neighbor = housingData %>% group_by (Neighborhood) %>%
  summarise(meanmarketvalue=mean(Market.Value.per.SqFt),min=min(Market.Value.per.SqFt)
            ,max=max(Market.Value.per.SqFt),oldestyear=min(Year.Built)
            ,newestyear=max(Year.Built))
housingData.Boro =housingData %>% group_by (Boro) %>%
  summarise(meanmarketvalue=mean(Market.Value.per.SqFt),min=min(Market.Value.per.SqFt)
            ,max=max(Market.Value.per.SqFt),oldestyear=min(Year.Built)
            ,newestyear=max(Year.Built))
housingData.neighbor
housingData.Boro

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
ggplot(housingData,aes(y=Year.Built,x=Boro,col=Boro)) + geom_boxplot()
ggplot(housingData,aes(y=Market.Value.per.SqFt,x=Boro,col=Boro)) + geom_boxplot()

plot(housingData$Market.Value.per.SqFt,housingData$Year.Built,col=as.factor(housingData$Boro),pch=19)
legend("bottomright",legend=unique(housingData$Boro),fill=as.factor(unique(housingData$Boro)))
plot(housingData$Market.Value.per.SqFt,housingData$Year.Built,col=as.factor(housingData$Neighborhood),pch=19)
cor(housingData$Market.Value.per.SqFt,housingData$Year.Built)

```

c. Write a summary about your findings from this exercise.
From these graphs above, the market value per square feet in Manhattan is the largest among 5 places: Bronx, Brooklyn, Manhattan, Queens and Staten Island.
Additionally, the correlation between market value per square feet and the year built is -0.0955, which is really close to 0, which means that there seems to be really small relationship between the year house was built and the market value per square feet.


